### 2021-06-20 comparison.Rmd

---
title: "Comparison"
output: html_document
---

```{r}
library(ggplot2)
library(ggpubr)
library(kableExtra)
Kalmanvarplot<-ggplot(DLM.df,aes(x=timeframe))+
  geom_line(aes(y=C))+
  labs(x="Time",
       y="")+
  ggtitle("Kalman Filter")+
  theme_bw()

Varplot<-function(dataframe,fun,title){

Filt<-Filtervalues(fun)
mean<-Filt$mean
sd<-Filt$sd
dataframe<-data.frame(dataframe,mean,sd)

ggplot(dataframe,aes(x=timeframe))+
geom_line(aes(y=sd))+
  labs(x="Time",
       y="")+
  ggtitle(title)+
  theme_bw()

}

Varsisplot<-Varplot(dt,sis,"SIS")
Varpfplot<-Varplot(dt,pf,"PF")
Varapfplot<-Varplot(dt,pf,"APF")
Varlwplot<-Varplot(dt,pf,"LW")

ggarrange(Kalmanvarplot,Varsisplot,Varpfplot,Varapfplot,Varlwplot)

```

```{r}
truex<-x
estimatedx<-matrix(NA,ncol=5,nrow=3)
colnames(estimatedx)<-c("N","KF","PF","APF","LWF")
estimatedx[,1]<-c(20,100,500)
RMSE<-function(x,xhat){sqrt(mean((x - xhat)^2))}

i=1
for(N in c(20,100,500)){

DLM1<-DLM(y,sigma2,tau2,m0,C0)
pf1<-PFfun(y,N,m0,C0,tau,sigma)
apf1<-APFfun(y,N,m0,C0,tau,sigma)
lwf1<-LWfun(y,N,m0,C0,1,1,1,1,delta,unif=T)

estimatedx[i,2]<-RMSE(truex,DLM1$m)
estimatedx[i,3]<-RMSE(truex,Filtervalues(pf1)$mean)
estimatedx[i,4]<-RMSE(truex,Filtervalues(apf1)$mean)
estimatedx[i,5]<-RMSE(truex,Filtervalues(lwf1)$mean)
i=i+1

}

```

```{r echo=FALSE, message=FALSE, warning=FALSE, results="asis"}
kbl(estimatedx, align="c", longtable = T, booktabs = T, caption="RMSE",digits=3)
```